\chapter{Transformative Tablets: Pioneering Success for Visually Impaired Students Through Innovative Apps}\label{ios-devices}

In an era where technology shapes the landscape of education, tablets have emerged as transformative tools, providing visually impaired students with unprecedented access to knowledge and fostering independence in their academic journeys. Both iPad and Android devices offer user-friendly interfaces and a diverse array of applications specifically tailored to bridge the accessibility gap. This chapter explores how tablets, in tandem with purpose-built apps, are not just tools but catalysts for success in the educational journey of visually impaired students.
\cite{OrientationMobilityInstruction}

Tablets serve as dynamic portals for visually impaired learners, offering a multi-sensory approach to engagement. Their unique functionalities, combined with a robust ecosystem of accessibility apps, empower students to navigate the digital realm with confidence and independence.

\section{Tablet Considerations}\label{tablet-considerations}

When selecting a tablet for students with visual impairments, consider the following:

\begin{itemize}
    \item \emph{Accessibility features:} Ensure compatibility with screen readers and magnification tools (e.g., VoiceOver for iOS, TalkBack for Android).
    \cite{AndroidAccessibility}
    \item \emph{Tactile features, size, and weight:} Choose a device that accommodates the student's specific needs.
    \item \emph{Contrast and color settings:} High contrast and customizable color settings, as well as text-to-speech functionalities, enhance readability.
    \item \emph{Screen size:} Larger screens can help reduce visual fatigue and improve usability, but balance with portability.
    \item \emph{App compatibility:} Ensure the device supports a variety of educational and accessibility apps.
    \item \emph{Visual fatigue:} Avoid prioritizing brightness; instead, focus on resolution and screen area. Adjust luminance for students with photophobia.
    \item \emph{AI Integration:} Modern tablets now offer enhanced AI-powered accessibility features, including improved object recognition and natural language processing.
\end{itemize}

Contrast ratio is especially important for students with visual impairments. Understanding and prioritizing contrast ratio helps foster an inclusive and enriching educational environment.

\section{Tablet Options}\label{tab:tablet-options}
When choosing an Android Tablet or iPad for a student with visual impairments, several factors must be considered to ensure that the student receives free and appropriate public education. The first factor to consider is the screen contrast ratio. A high contrast ratio is essential for students with visual impairments as it makes it easier for them to read text and view images on the screen. For Android Tablets, the W3C recommends a contrast ratio of at least 4.5:1 for small text and 3.0:1 for large text.
\cite{GoogleColorContrast}
Apple devices continue to excel with their "Increase Contrast" feature and additional accessibility enhancements.
\cite{iMoreContrast}

The second factor to consider is the size of the screen. A larger screen is beneficial for students with visual impairments as it allows them to view text and images more clearly. Tablets usually have larger screens than smartphones, making them a better choice for students with visual impairments\footnote{\raggedright \href{https://www.afb.org/blindness-and-low-vision/using-technology/cell-phones-tablets-mobile/smartphone-or-tablet-which}{American Foundation for the Blind. (n.d.). Smartphone or Tablet: Which is Best for You? Retrieved December 19, 2023}}. However, it is important to note that larger screens come at the expense of portability. Therefore, it is essential to find a balance between screen size and portability.

The third factor to consider is the availability of accessible apps. Both Android and iOS devices have built-in accessibility features such as screen readers, magnifiers, and high contrast modes.
\cite{AAOApps}
\cite{AFBiOS}
Additionally, there are several apps available that are specifically designed for students with visual impairments. Enhanced AI-powered apps like Be My Eyes now include Be My AI functionality, while SeeingAI continues to evolve with improved object recognition capabilities.
\cite{AAOTechnologyTools}

\emph{Tables \ref{tab:android-tablets} through \ref{tab:chromeOS-tablets}} describe current tablet computers that are available for students with visual impairments.

\subsection{AndroidOS 14+ Tablets}
\tagpdfsetup{table/header-rows={1}}
\begin{longtblr}[
  caption = {AndroidOS 14+ tablets suitable for students with visual impairments (Updated 2025)},
  label = {tab:android-tablets},
  remark{Note} = {Summary: Comprehensive list of Android tablets running OS 14 or higher, showing model name and screen size in inches}
]{
  colspec = {X[l] X[l]},
  rowhead = 1,
  hlines,
  stretch = 1.5
}
Tablet & Screen Size \\
Google Pixel Tablet & 10.9 \\
Samsung Galaxy Tab A9 & 8.7 \\
Samsung Galaxy Tab A9+ & 11.0 \\
Samsung Galaxy Tab S9 & 11.0 \\
Samsung Galaxy Tab S9 FE & 10.9 \\
Samsung Galaxy Tab S9 FE+ & 12.4 \\
Samsung Galaxy Tab S9 Ultra & 14.6 \\
Samsung Galaxy Tab S9+ & 12.4 \\
Samsung Galaxy Tab S10+ & 12.4 \\
Samsung Galaxy Tab S10 Ultra & 14.6 \\
OnePlus Pad & 11.6 \\
OnePlus Pad 2 & 12.1 \\
Lenovo Tab P12 & 12.7 \\
Lenovo Tab P12 Pro & 12.6 \\
Lenovo Tab M10 Plus (4th Gen) & 10.6 \\
Lenovo Tab M11 & 11.0 \\
Xiaomi Pad 6 & 11.0 \\
Xiaomi Pad 6 Pro & 11.0 \\
Xiaomi Pad 6S Pro & 12.4 \\
Honor Pad 9 & 12.1 \\
Realme Pad 2 & 11.5 \\
Redmi Pad Pro & 12.1 \\
Oppo Pad Air 2 & 11.4 \\
Vivo Pad 3 & 12.1 \\
\end{longtblr}

\subsection{iPadOS Tablets}
\tagpdfsetup{table/header-rows={1}}
\begin{longtblr}[
  caption = {iPadOS tablets suitable for students with visual impairments (Updated 2025)},
  label = {tab:ios-tablets},
  note = {Current Apple iPad models with their costs and screen sizes, arranged by model type. Prices as of July 2025}
]{
  colspec = {X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\bfseries},
  hlines,
  stretch = 1.5
}
Tablet & Cost & Screen Size \\
Apple iPad 10.9 (10th Gen) & \$349 & 10.9 \\
Apple iPad Air 11 (M2) & \$599 & 11.0 \\
Apple iPad Air 13 (M2) & \$799 & 13.0 \\
Apple iPad Pro 11 (M4) & \$999 & 11.0 \\
Apple iPad Pro 13 (M4) & \$1299 & 13.0 \\
Apple iPad mini 7 (A17 Pro) & \$499 & 8.3 \\
\end{longtblr}

\subsection{Windows OS Tablets}
\tagpdfsetup{table/header-rows={1}}
\centering
\begin{longtblr}[
  caption = {Windows OS tablets suitable for students with visual impairments (Updated 2025)},
  label = {tab:windows-tablets},
  note = {Windows-based tablets and 2-in-1 devices suitable for visually impaired students, listing model, price, and display size}
]{
  colspec = {X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\bfseries},
  hlines,
  stretch = 1.5
}
Tablet & Cost & Screen Size \\
Microsoft Surface Pro 11 & \$1199 & 13.0 \\
Microsoft Surface Pro 10 & \$1099 & 13.0 \\
Microsoft Surface Go 4 & \$629 & 10.5 \\
Microsoft Surface Laptop Studio 2 & \$1999 & 14.4 \\
Dell XPS 13 2-in-1 & \$1299 & 13.0 \\
HP Spectre x360 & \$1149 & 13.5 \\
Lenovo ThinkPad X1 Tablet & \$1449 & 13.0 \\
Asus VivoBook 13 Slate & \$849 & 13.3 \\
\end{longtblr}

\subsection{ChromeOS Tablets}
\tagpdfsetup{table/header-rows={1}}
\begin{longtblr}[
  caption = {ChromeOS tablets suitable for students with visual impairments (Updated 2025)},
  label = {tab:chromeOS-tablets},
  note = {Available ChromeOS tablets and their specifications, focusing on cost-effective options for educational use}
]{
  colspec = {X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\bfseries},
  hlines,
  stretch = 1.5
}
Tablet & Cost & Screen Size \\
Lenovo Chromebook Duet 3 & \$349 & 11.0 \\
Lenovo Chromebook Duet 5 & \$449 & 13.3 \\
HP Chromebook x2 11 & \$599 & 11.0 \\
Asus Chromebook CM3 & \$399 & 10.5 \\
Acer Chromebook Spin 714 & \$729 & 14.0 \\
\end{longtblr}

\section{Mobile Applications}\label{tab:tablelet-apps}
Mobile apps run on tablets are becoming increasingly important for students with visual impairments to access a free and appropriate public education. These apps can provide students with access to digital content, assistive technology, and other tools that can help them succeed in their studies. High-quality mobile apps can help students with visual impairments access the same educational materials as their sighted peers and participate fully in the curriculum. They can also help improve literacy skills, comprehension, and productivity. In this section, we will explore the importance of high-quality mobile apps for students with visual impairments and discuss some of the best apps available on the market today.

\emph{Accessibility Training/Auditory Games}
\begin{longtblr}[
  caption = {Mobile apps for accessibility training and auditory games for students with visual impairments (Updated 2025)},
  label = {tab:chapter2:accessibility-training-apps},
  note = {Educational apps designed to teach screen reader gestures and provide auditory game experiences, including current pricing and platform availability}
]{
  colspec = {X[l] X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\normalfont},
  hlines,
  stretch = 1.5
}
App & Cost & Function & OS \\
CosmoBally in Space & free & Train VoiceOver Gestures & iOS/iPadOS \\
Ballyland Magic Plus & \$4.99 & Train VoiceOver Gestures & iOS/iPadOS \\
Ballyland Rotor & \$3.99 & Train VoiceOver rotor & iOS/iPadOS \\
Ballyland Stay Still Squeaky! & \$3.99 & Train VoiceOver Gestures & iOS/iPadOS \\
Blindfold Games Launcher & free\footnote{\raggedright Games purchased separately} & Sonic Games & iOS/iPadOS \\
Blindfold Tap and Swipe & free & Train VoiceOver Gestures & iOS/iPadOS \\
ObjectiveEd Games & free\footnotemark[\value{footnote}] & Sonic Games & iOS/iPadOS \\
VO Lab & \$5.99 & Train VoiceOver Gestures & iOS/iPadOS \\
Screenreader & free & Train Accessibility Gestures & iOS/iPadOS, Android 14+ \\
\end{longtblr}

\subsection{Cortical Vision Impairment}
\begin{longtblr}[
  caption = {Mobile apps for cortical vision impairment (CVI) training for students with visual impairments (Updated 2025)},
  remark{Note} = {Summary: Specialized apps designed for CVI training and assessment, including current pricing and supported platforms},
  label = {tab:chapter2:cvi-training-apps}
]{
  colspec = {X[l] X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\normalfont},
  hlines,
  stretch = 1.5
}
App & Cost & Function & OS \\
Art of Glow & free & CVI-based Vision Training & iOS/iPadOS \\
Big Band Patterns & \$39.99 & CVI-based Vision Training & iOS/iPadOS \\
Big Bang Pictures & \$39.99 & CVI-based Vision Training & iOS/iPadOS \\
CVI Connect & \$12/mo & CVI-based Vision Training & iOS/iPadOS \\
CVI Connect Pro & free\footnote{\raggedright Annual Price Per Enrolled Student: 1-5=\$350\quad6-10=\$300\quad11-15=\$250\quad16-19=\$200} & CVI-based Vision Training & iOS/iPadOS \\
CVI Toddler Visual Eye Train & free & CVI-based Vision Training & iOS/iPadOS \\
CVI Training (Color) & free & CVI-based Vision Training & iOS/iPadOS \\
CVI Training (Human face) & free & CVI-based Vision Training & iOS/iPadOS \\
CVI Training (Pattern) & free & CVI-based Vision Training & iOS/iPadOS \\
CVI Training (Recognition) & free & CVI-based Vision Training & iOS/iPadOS \\
CVI Training (Visual Tracking) & free & CVI-based Vision Training & iOS/iPadOS \\
Dexteria VMI & \$6.99 & CVI-based Vision Training & iOS/iPadOS \\
EDA Play & \$5.99 & CVI-based Vision Training & iOS/iPadOS \\
EDA Play ELIS & \$3.99 & CVI-based Vision Training & iOS/iPadOS \\
EDA Play PAULI & \$3.99 & CVI-based Vision Training & iOS/iPadOS \\
EDA Play TOBY & free & CVI-based Vision Training & iOS/iPadOS \\
EDA Play TOM & free & CVI-based Vision Training & iOS/iPadOS \\
EyeMove & free & CVI-based Vision Training & iOS/iPadOS \\
Little Bear Sees & \$5.99 & CVI-based Vision Training & iOS/iPadOS \\
Peekaboo Barn & \$3.99 & CVI-based Vision Training & iOS/iPadOS \\
Sensory Light Box & \$4.99 & CVI-based Vision Training & iOS/iPadOS \\
Tap-n-See Now & \$3.99 & CVI-based Vision Training & iOS/iPadOS \\
Visual Attention Therapy Lite & free & CVI-based Vision Training & iOS/iPadOS \\
\end{longtblr}

\subsection{Audiobook/Reading}
\begin{longtblr}[
  caption = {Mobile apps for audiobook, e-book, and DAISY reading for students with visual impairments (Updated 2025)},
  label = {tab:chapter2:audiobook-reading-apps}
]{
  colspec = {X[l] X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\normalfont},
  hlines,
  stretch = 1.5
}
App & Cost & Function & OS \\
Audible & free\footnote{\raggedright requires books to be purchased or subscription \$14.95/mo} & Audiobook & iOS/iPadOS, AndroidOS 14+ \\
BARD Mobile & free\footnote{\raggedright requires account with local affiliate State Library for the Blind} & e-Book & iOS/iPadOS, AndroidOS 14+ \\
Bookshare Reader & free & DAISY Reader & iOS/iPadOS \\
Dolphin EasyReader & free & DAISY Reader & iOS/iPadOS, AndroidOS 14+ \\
KNFB Reader (OneStepReader) & \$109.99 & OCR/Reading & iOS/iPadOS, AndroidOS 14+ \\
Kindle & free\footnote{\raggedright requires books to be purchased or Kindle Unlimited \$11.99/mo} & e-Book & iOS/iPadOS, AndroidOS 14+ \\
Libby & free\footnote{\raggedright requires library card from participating library} & Audiobook & iOS/iPadOS, AndroidOS 14+ \\
Speech Central & free\footnote{\raggedright All paid features free for VoiceOver users} & Text-to-Speech & iOS/iPadOS, AndroidOS 14+ \\
VoiceDream Reader & \$19.99\footnote{\raggedright Premium voices available for additional cost} & DAISY Reader & iOS/iPadOS \\
\end{longtblr}

\emph{Productivity/Schoolwork/Optical Character Recognition}
\begin{longtblr}[
  caption = {Mobile apps for productivity, schoolwork, and optical character recognition (OCR) for students with visual impairments (Updated 2025)},
  label = {tab:chapter2:productivity-ocr-apps}
]{
  colspec = {X[l] X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\normalfont},
  hlines,
  stretch = 1.5
}
App & Cost & Function & OS \\
Aiko & free & AI Speech to text & iOS/iPadOS \\
Ballyland Code 1: Say Hello & \$3.99 & Auditory Coding & iOS/iPadOS \\
Ballyland Code 2: Give Rotor & \$3.99 & Auditory Coding & iOS/iPadOS \\
Ballyland Code 3: Pick Up & \$3.99 & Auditory Coding & iOS/iPadOS \\
Clusiv & free\footnote{\raggedright Full access provided through Vocational Rehabilitation} & Online learning platform & iOS/iPadOS \\
Code Quest & free & Auditory Coding & iOS/iPadOS \\
Desmos Graphing Calculator & free & Accessible Graphing & iOS/iPadOS, AndroidOS 14+ \\
Desmos Scientific Calculator & free & Accessible Scientific Calculator & iOS/iPadOS, AndroidOS 14+ \\
Envision AI & free\footnote{\raggedright Premium features \$9.99/mo} & AI-powered OCR & iOS/iPadOS, AndroidOS 14+ \\
GoodNotes 5 & \$7.99 & Scan \& Markup Documents & iOS/iPadOS, AndroidOS 14+ \\
Microsoft 365 & free\footnote{\raggedright Full features with subscription \$6.99/mo} & Office Suite & iOS/iPadOS, AndroidOS 14+ \\
Notability & \$11.99 & Scan \& Markup Documents & iOS/iPadOS \\
SeeingAI & free & Talking Camera & iOS/iPadOS, AndroidOS 14+ \\
TapTapSee & free & Talking Camera & iOS/iPadOS, AndroidOS 14+ \\
Voice Control for ChatGPT & free\footnote{\raggedright Premium features \$4.99/mo} & AI Assistant & iOS/iPadOS, AndroidOS 14+ \\
\end{longtblr}

\emph{Orientation \& Mobility / Navigation}
\begin{longtblr}[
  caption = {Mobile apps for orientation, mobility, and navigation for students with visual impairments (Updated 2025)},
  label = {tab:chapter2:navigation-apps}
]{
  colspec = {X[l] X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\normalfont},
  hlines,
  stretch = 1.5
}
App & Cost & Function & OS \\
Apple Maps & free & Turn by Turn Navigation & iOS/iPadOS \\
Be My Eyes & free & Visual assistance via volunteers & iOS/iPadOS, AndroidOS 14+ \\
BlindSquare & \$49.99 & GPS Navigation & iOS/iPadOS \\
Clew & free & Indoor navigation & iOS/iPadOS, AndroidOS 14+ \\
GoodMaps Explore & free & Indoor/Outdoor navigation & iOS/iPadOS \\
Google Maps & free & Turn by Turn Navigation & iOS/iPadOS, AndroidOS 14+ \\
Lazarillo & free & GPS navigation & iOS/iPadOS, AndroidOS 14+ \\
Moovit & free\footnote{\raggedright Premium features \$3.99/mo} & Public Transit & iOS/iPadOS, AndroidOS 14+ \\
Oko & free & Smart traffic detection & iOS/iPadOS, AndroidOS 14+ \\
Soundscape & free & 3D audio navigation & iOS/iPadOS, AndroidOS 14+ \\
Waymap & free & Indoor/Outdoor navigation & iOS/iPadOS, AndroidOS 14+ \\
\end{longtblr}

\subsection{Independent Living Skills}
\begin{longtblr}[
  caption = {Mobile apps for independent living skills for students with visual impairments (Updated 2025)},
  label = {tab:chapter2:independent-living-apps}
]{
  colspec = {X[l] X[l] X[l] X[l]},
  rowhead = 1,
  row{1} = {font=\normalfont},
  hlines,
  stretch = 1.5
}
App & Cost & Function & OS \\
CashReader & free\footnote{\raggedright Premium features \$2.99/mo, \$19.99/yr, or \$49.99/lifetime} & Currency identification & iOS/iPadOS, AndroidOS 14+ \\
Lookout & free & AI-powered assistance & AndroidOS 14+ \\
Magnifier & free & Built-in magnification & iOS/iPadOS \\
Menus4All & free\footnote{\raggedright Premium subscription \$3.99/mo or \$39.99/yr} & Restaurant menus & iOS/iPadOS \\
PenFriend & free & Labeling system app & iOS/iPadOS \\
Supersense & free\footnote{\raggedright Premium features \$9.99/mo} & AI scene description & iOS/iPadOS, AndroidOS 14+ \\
Voice Assistant & free & Voice control & iOS/iPadOS, AndroidOS 14+ \\
\end{longtblr}

\section{Conclusion}
\label{sec:conclusion-tablets}
The landscape of assistive technology for visually impaired students continues to evolve rapidly, with new applications and enhanced features being released regularly. When selecting tablets and applications for educational use, it is essential to consider the specific needs of each student, the available budget, and the compatibility with existing assistive technologies. Regular updates to both hardware and software ensure that students have access to the most current and effective tools for their educational success.

The integration of artificial intelligence and machine learning in accessibility applications has significantly improved the user experience for visually impaired students. From enhanced object recognition to more accurate optical character recognition, these technological advances continue to break down barriers and provide greater independence in educational settings.

\begin{thebibliography}{99}
\bibitem{OrientationMobilityInstruction} Orientation \& Mobility Instruction. Retrieved December 19, 2023.
\bibitem{AndroidAccessibility} Android Accessibility Features. Retrieved December 19, 2023.
\bibitem{GoogleColorContrast} Google. (n.d.). Color contrast - Android Accessibility Help. Retrieved December 19, 2023.
\bibitem{iMoreContrast} iMore. (n.d.). How to increase contrast for visual accessibility on iPhone and iPad. Retrieved December 19, 2023.
\bibitem{AAOApps} American Academy of Ophthalmology. (n.d.). 30 Apps, Devices and Technologies for People With Vision Impairments. Retrieved December 19, 2023.
\bibitem{AFBiOS} American Foundation for the Blind. (n.d.). Apple iOS for iPhone and iPad: Considerations for Users with Visual Impairments. Retrieved December 19, 2023.
\bibitem{AAOTechnologyTools} American Academy of Ophthalmology. (n.d.). Technology Tools for Children with Low Vision. Retrieved December 19, 2023.
\end{thebibliography}
