\hypertarget{low-vision}{}\chapter[\raggedright Amplifying Vision:\hfill\break The Vital Role of Video Magnification Products in Fostering Literacy\hfill\break and Success for Visually Impaired Students]{Amplifying Vision: The Vital Role of Video Magnification Products in Fostering Literacy and Success for Visually Impaired Students}\label{low-vision}
\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}
{\let\clearpage\relax\localtableofcontents\let\clearpage\relax\locallistoftables}\newpage
\extramarks{Vision Department Technology Needs}{Chapter 6: Amplifying Vision}
In the realm of visual impairment, the quest for literacy and academic success is a journey characterized by innovation and adaptability. For visually impaired students, the challenge of accessing printed materials, charts, and visual content is met with a powerful solution—video magnification products. The indispensable role that video magnification plays in providing enhanced visual access, breaking down barriers to literacy, and empowering students to navigate the educational landscape with confidence cannot be overstated.

The significance of video magnification products lies in their ability to transform the visual experience for students with visual impairments. As we navigate this chapter, we will unravel the sophisticated functionalities of these devices, showcasing how they go beyond traditional magnification methods to provide an immersive and dynamic visual experience. Whether exploring the pages of a textbook, deciphering intricate diagrams, or engaging with digital content, video magnification stands as a technological ally, ensuring that every student can access and interpret visual information with ease.

In the pursuit of literacy, the role of video magnification becomes increasingly pivotal, particularly in subjects where visual content is integral to comprehension. This chapter will delve into how these products facilitate not only enhanced readability but also active participation in classroom discussions, visual learning activities, and the overall educational experience. By providing visually impaired students with a clear and magnified view of the visual world, video magnification products serve as gateways to knowledge, fostering a sense of inclusion and leveling the playing field in academic settings.

It is evident that these tools are not mere aids; they are essential components in the arsenal of resources necessary for the success of visually impaired students. Video magnification imperatively contributes to shaping a learning environment where visual content is accessible to all, ensuring that literacy and success are attainable goals for every student, regardless of their visual abilities.


\pagebreak\hypertarget{video-magnification-devices}{}\section{Video Magnification
 Devices}\label{video-magnification-devices}
When purchasing electronic portable magnifiers for students with visual impairments, it is important to consider the following factors to ensure that they can access a free and appropriate public education\footnote{\raggedright \textit{cf}., \href{http://www.perkins.org/resource/choosing-appropriate-video-magnifier/}{Perkins School for the Blind. (n.d.). Choosing an appropriate video magnifier. Retrieved December 19, 2023}}:
\begin{itemize}[leftmargin=*]
 \item \textbf{Magnification power}: The magnification power of the magnifier should be appropriate for the student’s needs. Some magnifiers have a fixed magnification, while others have adjustable magnification.
 \item \textbf{Portability}: Portable magnifiers are ideal for students who need to move around the classroom or school. They should be lightweight and easy to carry.
 Battery life: Battery life is an important consideration for portable magnifiers. The battery should last long enough to get through a school day without needing to be recharged.
 \item \textbf{Ease of use}: The magnifier should be easy to use and adjust. It should have large buttons and controls that are easy to locate and operate.
 Compatibility: The magnifier should be compatible with the student’s other assistive technology devices, such as screen readers and braille displays.
 \item \textbf{Cost}: The cost of the magnifier should be reasonable and within the school’s budget.
\end{itemize}
These considerations will help ensure that students with visual impairments have access to the tools they need to succeed in school. \textit{Table \ref{tab:table21}} lists current available video magnification devices for students with visual impairments.
 
\begin{longtable}[]{@{}
 >{\raggedright\arraybackslash}m{.30\textwidth}
 >{\raggedright\arraybackslash}m{.15\textwidth}
 >{\raggedright\arraybackslash}m{.3\textwidth}
 >{\raggedright\arraybackslash}b{.25\textwidth}@{}
 }
 \toprule
 
 \textbf{Model} & \textbf{Cost} & \textbf{Deployment} & \textbf{Company} \\
 \midrule
 \endhead \hline \\
 \multicolumn{4}{r}{\textbf{Continued on Next Page}} \endfoot
 \endlastfoot
 AceSight VR & \$2,695 & VR Headset & Zoomax \\ \cdashline{1-4}
 Acesight & \$4,295 & E-Glasses & Zoomax \\ \cdashline{1-4}
 Acesight 8 & \$2,995 & E-Glasses & Zoomax \\ \cdashline{1-4}
 Acuity 22 & \$2,695 & Desktop & Irie AT \\ \cdashline{1-4}
 Acuity 22 Speech & \$3,695 & Desktop & Irie AT \\ \cdashline{1-4}
 Amigo & \$1,400 & Portable & Enhanced Vision \\ \cdashline{1-4}
 Cloverbook Plus & \$2,295 & Mobile & Irie-AT \\ \cdashline{1-4}
 Cloverbook Pro & \$2,995 & Mobile & Irie-AT \\ \cdashline{1-4}
 Connect 12 & \$2,695 & Desktop\break Mobile & Humanware \\ \cdashline{1-4}
 Connect 12 (10x) & \$2,845 & Desktop\break Mobile & Humanware \\ \cdashline{1-4}
 Connect 12 (25x) & \$2,895 & Desktop\break Mobile & Humanware \\ \cdashline{1-4}
 Distance Camera & TBD & Hand-Held & Zoomax \\ \cdashline{1-4}
 explore 12 & \$1,895 & Desktop & Humanware \\ \cdashline{1-4}
 explore 5 & \$845 & Hand-Held (5'' screen) & Humanware \\ \cdashline{1-4}
 explore 8 & \$1,275 & Hand-Held (8'' screen) & Humanware \\ \cdashline{1-4}
 I-See 22'' & \$2,095 & Desktop & Irie AT \\ \cdashline{1-4}
 Juno & \$1,392 & Hand-Held (7'' screen) & APH \\ \cdashline{1-4}
 Jupiter Portable Magnifier & \$3,599.00 & Desktop\break Mobile, but heavy & APH \\ \cdashline{1-4}
 Luna 6 & \$795 & Hand-Held (6'' screen) & Zoomax \\ \cdashline{1-4}
 Luna 8 & \$895 & Hand-Held (8'' screen) & Zoomax \\ \cdashline{1-4}
 Luna Eye & TBD & Hand-Held & Zoomax \\ \cdashline{1-4}
 Luna HD Pro & \$2,995 & Desktop & Zoomax \\ \cdashline{1-4}
 Luna S & \$385 & Hand-Held (4.3'' screen) & Zoomax \\ \cdashline{1-4}
 MAGNA 3 & \$149 & Hand-Held (3.5'' screen) & Orbit Research \\ \cdashline{1-4}
 MAGNA 4 & \$199 & Hand-Held (4.3'' screen) & Orbit Research \\ \cdashline{1-4}
 MAGNA 5 & \$249 & Hand-Held (5'' screen) & Orbit Research \\ \cdashline{1-4}
 MATT Connect v2 & \$3,895 & Desktop\break Mobile, but heavy & APH \\ \cdashline{1-4}
 Magnibot & \$2,995 & Desktop\break Mobile & Trysight \\ \cdashline{1-4}
 Magnilink Air & \$5,995 & Desktop & Low Vision International \\ \cdashline{1-4}
 MagniLink Tab & \$5,895 & Desktop & Low Vision International \\ \cdashline{1-4}
 MagniLink One & \$2,395 & Desktop & Low Vision International \\ \cdashline{1-4}
 MagniLink S Premium & \$4,295 & Mobile & Low Vision International \\ \cdashline{1-4}
 MagniLink Vision & \$3,190 - \$4,250 & Desktop & Low Vision International \\ \cdashline{1-4}
 MagniLink WifiCam & \$3,695 & Mobile & Low Vision International \\ \cdashline{1-4}
 MagniLink Zip & \$3,625 & Desktop & Low Vision International \\ \cdashline{1-4}
 Merlin Mini & \$3,570 & Mobile & Enhanced Vision \\ \cdashline{1-4}
 ONYX Desk set HD & \$3,330 & Desktop & Freedom Scientific \\ \cdashline{1-4}
 ONYX OCR & \$4,520 & Desktop & Freedom Scientific \\ \cdashline{1-4}
 Panda HD & \$2,098 & Desktop & Zoomax \\ \cdashline{1-4}
 Pebble HD & \$656 & Handheld & Enhanced Vision \\ \cdashline{1-4}
 RUBY & \$600.60 & Hand-Held (4.3'' screen) & Freedom Scientific \\ \cdashline{1-4}
 RUBY 10 & \$1,640 & Hand-Held (10'' Screen) & Freedom Scientific \\ \cdashline{1-4}
 RUBY 7 HD & \$1,317.75 & Hand-Held (7'' Screen) & Freedom Scientific \\ \cdashline{1-4}
 RUBY HD & \$710.85 & Hand-Held (4.3'' screen) & Freedom Scientific \\ \cdashline{1-4}
 RUBY XL HD & \$987.00 & Hand-Held (5'' screen) & Freedom Scientific \\ \cdashline{1-4}
 Reveal 16 & \$3,295 & Desktop & Humanware \\ \cdashline{1-4}
 Reveal 16 (XY table) & \$3,995 & Desktop & Humanware \\ \cdashline{1-4}
 Reveal 16i & \$4,295 & Desktop & Humanware \\ \cdashline{1-4}
 Reveal 16i (XY table) & \$4,995 & Desktop & Humanware \\ \cdashline{1-4}
 Snow 12 & \$1,395 & Desktop\break Mobile & Zoomax \\ \cdashline{1-4}
 Snow Pad & TBD & Hand-Held & Zoomax \\ \cdashline{1-4}
 TOPAZ EZ HD & \$3,081.75 & Desktop & Freedom Scientific \\ \cdashline{1-4}
 TOPAZ OCR & \$4,640.00 & Desktop & Freedom Scientific \\ \cdashline{1-4}
 TOPAZ XL HD & \$4,045.00 & Desktop & Freedom Scientific \\ \cdashline{1-4}
 Tactonum Pro & £10,000 & Desktop\break Not Readily Mobile & Tactonum \\ \cdashline{1-4}
 Taxtonum Reader & £3,795.00 & Desktop\break Not Readily Mobile & Tactonum \\ \cdashline{1-4}
 Transformer HD & \$3,565 & Mobile & Enhanced Vision \\ \cdashline{1-4}
 Traveller HD & \$656 & Mobile & Optelec \\ \cdashline{1-4}
 \hline
 \caption{ Video Magnification Devices}\label{tab:table21}
\end{longtable}