\chapter{Conclusion}\label{main-conclusion}

\section{~~Technology as the Cornerstone of Educational Equity}
\label{sec:conclusion-technology-equity}

This document has demonstrated that technology\index{technology} is not merely an accommodation for students with visual impairments\index{visual impairment}—it is the cornerstone of educational equity\index{educational equity}. Achieving true equity requires more than compliance\index{accessibility!legal accessibility} with legal\index{accessibility!legal} mandates; it demands a proactive, evidence-based approach to selecting, implementing, and supporting technology that enables visually impaired students to access, participate in, and excel within the educational environment. \supercite{IntroductionChapter1, Burgstahler2015, Kelly2011}

\section{~~Key Findings and Imperatives}
\label{sec:conclusion-key-findings}

\begin{itemize}
	\item \emph{Hardware\index{hardware} and Performance\index{troubleshooting!performance}:} The responsiveness of assistive technologies\index{assistive technology} such as screen readers is fundamentally dependent on robust\index{accessibility!principles} hardware. Underpowered devices create unacceptable barriers to learning, while high-performance systems empower students to engage with digital content on equal footing with their sighted peers.\footnote{See Chapter 1: Impact of Hardware Limitations on Screen Reader Response Latency\index{latency}.}
	\item \emph{Accessible Devices and Materials:} The integration of tablets, braille notetakers\index{notetaker!braille notetaker}, refreshable braille displays\index{braille}, video magnifiers\index{video magnifier}, and high-quality embossers expands access to literacy, STEM\index{STEM}, and the broader curriculum. Tactile graphics\index{tactile graphics} and 3D printed models transform abstract concepts into tangible learning experiences.\footnote{See Chapters 2–5}
	      Tools\index{sonification!tools} such as text-to-speech\index{text-to-speech} engines, DAISY\index{DAISY} readers, accessible GPS\index{GPS}, and daily living technologies foster not only academic achievement but also independence\index{independence}, self-advocacy, and community participation. \cite{Chapters6to8}
	\item \emph{Assessment\index{assistive technology!assessment} and Individualization:} There is no one-size-fits-all solution. Individualized assessment frameworks, such as the SETT model, and ongoing collaboration among students, families, and educators are essential to matching technology\index{technology} to each learner's unique needs.\footnote{See Appendix 3}
	      Effective technology use is sustained by high-quality training, troubleshooting\index{troubleshooting} resources\index{3D printing!resources}, and a commitment to continuous improvement. The appendices provide practical guidance for educators and families to ensure ongoing success. \cite{Appendices1to4}
	      Attention to accessible fonts, formatting, and instructional materials\index{instructional materials} ensures that all students—regardless of their degree of vision—can engage with content in a way that maximizes comprehension and comfort. \cite{Appendix5}
\end{itemize}

\section{~~A Call to Action}
\label{sec:conclusion-call-to-action}

Educational equity\index{educational equity} for students with visual impairments\index{visual impairment} is not achieved by chance, but by design. It is the result of intentional decisions to prioritize accessibility\index{accessibility}, invest in appropriate technology, and foster a culture of inclusion and high expectations. As technology continues to evolve, so too must our commitment to leveraging it as a force for justice and opportunity.

By embracing technology as a driver of equity, we empower visually impaired students to pursue academic excellence, independence\index{independence}, and full participation in school and society. The tools\index{sonification!tools}, strategies, and frameworkspresented in this document are not merely recommendations—they are a roadmap for transforming educational outcomes and realizing the promise of equity for all learners.

Let us move forward with resolve, ensuring that every student—regardless of vision—has the technology\index{technology}, support\index{troubleshooting!support}, and opportunity to thrive.